---
title: "Using oLlama in R"
author: "Luis Sattelmayer"
date: "2025-06-09"
format: html
---

This document demonstrates how to use the `rollama` package in R to interact with the oLlama framework. This tutorial is very very strongly inspired by the one of [Johannes B. Gruber](https://jbgruber.github.io/rollama/articles/annotation.html#the-make_query-helper-function) who is also the author of the `rollama` package. 

```{r}
needs(tidyverse, rollama)
```

Before we can get started, you will have to install `ollama`. You can do so by following the instructions on the [ollama website](https://ollama.com/download). It will ask you to download a setup file which will then guide you through the local installation. The last step is to run a command in your terminal. 

Once you have done this, you should be able to run the following command to see whether the rollama package can successfully connect to the local ollama installation. 

```{r}
rollama::ping_ollama()
```



Before delving into the details on how to use the models in R, you can also check out the *ollama* package through your Terminal. If the installation was succesful, you can run "ollama"  to see the available commands that come with the ollama software. Here are some of the more important ones: 

- `ollama pull` to load a package
- `ollama ls` to see the models you have downloaded
- `ollama run` to run a model
- `ollama rm [insert the model name here]` to delete a model

The next step is to (down-)load a model. In this example, we will follow Gruber's model choice of `llama3.2:3b-instruct-q8_0`. This is one of the smaller model that is well-suited for quick tests and demonstrations but should also already be able to handle more complex tasks. There are a variety of different models available, which you can find on the [ollama model page](https://ollama.com/search). They differ in size, capabilities, the type of tasks you can use them for and (quite importantly) the developing company. You will download them locally, which implies that you will need enough disk space for the model you choose. The smallest ones are around 1 to 3 GB, while the largest ones can be much bigger. Technically, the larger models should be able to handle more complex tasks, but they are also considerably slower.



```{r}
options(rollama_model = "llama3.2:3b-instruct-q8_0")
```


```{r}
econ <- read_csv("https://raw.githubusercontent.com/luissattelmayer/intro-css/refs/heads/main/data/tweets_annotations.csv") |> 
    slice(1:5)

econ
```

The `rollama` package provides a convenient way to create queries for the models. The `make_query()` function allows you to specify the text you want to analyze, the prompt for the model, and any additional system instructions. 

The `text` argument specifies the text you want to analyze, which in this case is the `text` column of the `econ` data frame. The `prompt` argument is a string that will be used as a prompt for the model. Here I simply tell the model what output categories I want as labels. The `template` argument allows you to specify how the input should be formatted. Here this means that the model will first see the prefix, followed by the text, after which we insert a paragraph and then the prompt. The `system` argument provides the actual instructions to the model, and the `prefix` argument specifies a prefix for the text to classify.

You can see that I am not giving the machine any examples of what the categories mean or what some ground-truth examples would look like. This here is purely a zero-shot classification task, which means that the model will have to rely on its pre-trained knowledge to classify the tweets.

```{r}
econ_queries <- make_query(
  text = econ$text,
  prompt = "Categories: 0, 1",
  template = "{prefix}{text}\n{prompt}",
  system = "Classify this tweet based on whether it speaks about economic issues or not. Don't give any more comments.",
  prefix = "You are a helpful and concise assistant that classifies tweets into two categories: economic issues (1) and non-economic issues (0). Please classify the following tweet:\n\n"
)
```


Let's now create a new column in the `econ` data frame that contains the model's annotations for each tweet. The `query()` function will send the custom `econ_queries` task to the model and return the results. The `verbose = TRUE` argument will print out the progress of the queries, and `output = "text"` specifies that we want the output in text format. The `model_params` argument allows us to set specific parameters for the model, such as the `seed` for reproducibility and the `temperature` for controlling randomness in the output. Setting a consistent seed, and a temperature of 0 will ensure that the model's output is deterministic, meaning that it will produce the same results each time you run the code. Great for reproducibility!

```{r}
econ$annotation <- query(econ_queries, verbose = TRUE, output = "text",  model_params = list(seed = 2507, temperature = 0))
econ
```

```{r}
# calculate F1 scores
MLmetrics::F1_Score(y_pred = econ$annotation, y_true = econ$label)
```

If you wanted to put this into a tidy workflow...

```{r}
econ_full <- read_csv("https://raw.githubusercontent.com/luissattelmayer/intro-css/refs/heads/main/data/tweets_annotations.csv")

econ_llama_annotated <- econ_full |>
  mutate(
    annotation = make_query(
      text = text,
      prompt = "Categories: 0, 1",
      template = "{prefix}{text}\n{prompt}",
      system = "Classify this tweet based on whether it speaks about economic issues or not. Don't give any more comments.",
  prefix = "You are a helpful and concise assistant that classifies tweets into two categories: economic issues (1) and non-economic issues (0). Please classify the following tweet:\n\n"
    ) |>
      query(screen = FALSE, output = "text", model_params = list(seed = 2507, temperature = 0))
  )

write_rds(econ_llama_annotated, "data/econ_llama_annotated.rds")

econ_llama_annotated
```

```{r}
MLmetrics::F1_Score(y_pred = econ_llama_annotated$annotation, y_true = econ_llama_annotated$label)
```


